---
title: "Homework 1"
author: "Ziang Zhang"
date: "2026-09-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

The `penguins` dataset contains measurements of Adelie, Chinstrap, and Gentoo penguins from the Palmer Archipelago. Important variables include `species` (Adelie, Chinstrap, or Gentoo), `island` (Biscoe, Dream, or Torgersen), `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`. The `sex` variable records female or male.

```{r load-penguins, message=FALSE}
library(dplyr)
library(ggplot2)
library(palmerpenguins)

data("penguins", package = "palmerpenguins")
```

The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r penguin-scatterplot}
penguin_scatterplot <-
  ggplot(
    penguins,
    aes(x = bill_length_mm, y = flipper_length_mm, color = species)
  ) +
  geom_point(na.rm = TRUE)

penguin_scatterplot

ggsave(
  filename = "penguin_scatterplot.png",
  plot = penguin_scatterplot,
  width = 6,
  height = 4
)
```

## Problem 2

```{r create-data}
set.seed(123)

normal_sample <- rnorm(10)

homework_data <-
  data.frame(
    normal_sample = normal_sample,
    is_positive = normal_sample > 0,
    character_variable = c(
      "red", "blue", "green", "red", "blue",
      "green", "red", "blue", "green", "red"
    ),
    factor_variable = factor(
      c(
        "low", "medium", "high", "low", "medium",
        "high", "low", "medium", "high", "low"
      )
    )
  )

homework_data
```

I used `pull()` to extract each variable from the data frame before calculating its mean. The numeric variable has a numeric mean. R also calculates a mean for the logical variable by treating `TRUE` as 1 and `FALSE` as 0; this mean is the proportion of values that are `TRUE`. The character and factor variables do not have meaningful arithmetic means, so their calls to `mean()` return `NA` with a warning.

```{r calculate-means}
mean(pull(homework_data, normal_sample))
mean(pull(homework_data, is_positive))
mean(pull(homework_data, character_variable))
mean(pull(homework_data, factor_variable))
```

Converting a logical vector to numeric maps `FALSE` to 0 and `TRUE` to 1. Converting these character values produces `NA` because they are not numeric strings. Converting a factor to numeric returns its internal level codes rather than the labels. This helps explain why the logical mean can be interpreted as a proportion, while character and factor means are not meaningful.

```{r numeric-conversions, results="hide", warning=FALSE}
as.numeric(pull(homework_data, is_positive))
as.numeric(pull(homework_data, character_variable))
as.numeric(pull(homework_data, factor_variable))
```